\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics

### 2017-09-20 journalSwarmControl/Introduction.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

%Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.  
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}. These activities require robots that behave intelligently.
Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire particle population.    
This paper examines object manipulation by a swarm of particles, each actuated by the same shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
The transportation methodology is similar to that in Sugawara's work \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.

>>>>

%Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.  
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}. These activities require robots that behave intelligently.
Limited computation and communication at small scales makes autonomous operation or direct control over individual robots difficult. 
Instead, this paper treats the robots as particles that are steered by a global control signal broadcast to the entire population.    
This paper examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
The transportation methodology is similar to that in Sugawara's work \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.

### 2017-09-20 journalSwarmControl/expResults.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

%Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source. 
%Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. 
%Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.  
This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all particles; {\bf convex-hull} draws a convex hull around the outermost particles; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse. Fig.~\ref{fig:Visualization} shows screenshots of the same particle swarm with each type of visual feedback. Full-state requires $2n$ data points for $n$ particles. Convex-hull requires at worst $2n$, but according to Har \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.  Mean requires two, and variance three, data points.  Mean and mean + variance are convenient even with millions of particles.

@@

Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.
Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.  As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].  All experiments lasting over 300s were removed, under the assumption that the user stopped playing. 
Using ANOVA analysis, we rejected the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
Anecdotal evidence from beta-testers who played the game suggests that tracking 100 particles is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator. Our beta-testers described convex-hull feedback as confusing and irritating.  A single particle left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm. Similarly, our algorithm must be robust to outliers.
%obscuring what the rest of the swarm is doing.

@@

\subsection{Varying noise}
%Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
This experiment varied the strength of %these 
disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{\rm{max}}$).

We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the particles move erratically.  Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise. This indicates swarm control is robust to IID noise.

>>>>

%Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source. 
%Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. 
%Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.  
This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all particles; {\bf convex-hull} draws a convex hull around the outermost particles; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse. Fig.~\ref{fig:Visualization} shows screenshots of the same particle swarm with each type of visual feedback. Full-state requires $2n$ data points for $n$ particles. Convex-hull requires at worst $2n$, but according to Har \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.  Mean requires two, and variance three, data points.  Because they do not increase with population size, mean and mean + variance are convenient even with millions of particles.

@@

Our hypothesis predicted a steady decrease in performance as the amount of visual feedback decreased.
Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.  As Fig.~\ref{fig:ResVaryVis}b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].  All experiments lasting over 300s were removed, under the assumption that the user stopped playing. 
Using ANOVA analysis, we rejected the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
Anecdotal evidence from beta-testers who played the game suggests that tracking 100 particles is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator. However, our beta-testers described convex-hull feedback as confusing and irritating since it is not robust to outliers.  A single particle left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
%obscuring what the rest of the swarm is doing.

@@

\subsection{Varying noise}
%Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
This experiment varied the strength of %these 
disturbances to study how noise (disturbance inputs) affects human control of large swarms. Noise was applied at every time step:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{\rm{max}}$).

We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the particles move erratically.  Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}a show only a 40\% increase in completion time for the maximum noise. This indicates swarm control is robust to IID noise.

### 2017-09-20 journalSwarmControl/experiment.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region.  To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.

@@

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j|x, \mathbf{u})].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| x_j,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
\text{end}& 
\end{align}
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach inspired by Spong \cite{spong2008robot} that attracts the swarm to the intermediate goal but repulses the swarm from in front of the object.
The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown in Fig.~\ref{fig:potentialField}b. These fields operate in place of $\mathbf{D}(\mathbf{b})$ only when the repulsive field is active.
\begin{align}
F_{\rm{att}} &= -\zeta \Delta \rho / \rho, \\
%\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right), \\
%\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
 F_{\rm{rep}} &=  \left\{
\begin{array}{ll}
      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\wedge~\phi <  \theta \\
      0, & \textrm{ otherwise} \\
\end{array},
\right.\\
F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}.
\end{align}

In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $\rho_0 = 3$. Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$. In both cases, $\Delta \rho$ is the position of the swarm.

>>>>

This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region. 
%We assume that particles pushing the object move slower than unencumbered particles, and that pushing with more particles increases the object speed
We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion.  Increasing the number of pushing particles increases the object speed.
%Although one particle might be able to move the object, it requires a long time for it to complete the task and probability of getting stuck when there is not enough particles is high.
 To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.

@@

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i|x, \mathbf{u})].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$j=1$ to $N$ do} \nonumber \\
&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| x_j,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
\text{end}& 
\end{align}
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach inspired by Spong \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object, as shown in Fig.~\ref{fig:potentialField}.
The repulsive potential field is centered at $\mathbf{b}$, the object's COM, and is active in a circular sector of angular width 2$\theta$ and radius $d_0$ aligned with $\mathbf{D}(\mathbf{b})$. $\mathbf{D}(\mathbf{b})$ is the desired direction of motion from~\eqref{eq:OptimalPolicy}.
\begin{align}
\mathbf{d} &=  [\bar{x},\bar{y}] - \mathbf{b}, \qquad   \phi =\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot  \mathbf{d}  }{\norm{\mathbf{D}(\mathbf{b})} \cdot \norm{\mathbf{d}} } \right), \nonumber \\ 
F_{\rm{att}} &= -\zeta \frac{\mathbf{d}  }{ \norm{\mathbf{d}}}, \nonumber\\
 F_{\rm{rep}} &=  \left\{
\begin{array}{ll}
      \eta( \frac{1}{\norm{\mathbf{d}}}- \frac{1}{d_0}) \frac{1}{\norm{\mathbf{d}}^2} \mathbf{d}, ~& \norm{\mathbf{d}} \leq d_0~\wedge~\phi <  \theta \\
      0, & \textrm{ otherwise} \nonumber \\
\end{array},
\right.\\
F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}, \nonumber\\
[x_{\rm{goal}}&, y_{\rm{goal}}] =  [\bar{x},\bar{y}]  + \frac{F_{\rm{pot}}}{\norm{F_{\rm{pot}}}}.
\end{align}
Here  $\eta$, $\zeta$, and $d_0$ are all positive parameters that scale the forces and $\norm{\mathbf{d}}$ is the distance from the swarm mean $ [\bar{x},\bar{y}]$ to the object COM.
In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $d_0 = 3$. Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $d_0 = 7.5$.

### 2017-09-20 journalSwarmControl/theory.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

Just as we used a PD controller in Eq.~\eqref{eq:PDcontrolPosition} to control the robots, a PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}), \nonumber\\
u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell_x/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.

@@

The mean and variance of the swarm cannot be controlled simultaneously, however if the dispersion due to Brownian motion is less than the maximum controlled speed, we can adopt the hybrid, hysteresis-based controller shown in Alg.~\ref{alg:MeanVarianceControl} to regulate the mean and variance.  Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds some $\sigma_{\rm{max}}^2$. 
 Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 
 This technique satisfies control objectives that evolve at different rates as Kloetzer's work ~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes. The process is illustrated in Fig.~\ref{fig:hysteresis}.

>>>>

Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean robot position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \label{eq:PDcontrolVariance}\\
u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell_x/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.  A similar argument applies to~\eqref{eq:PDcontrolVarianceIny}.

@@

The mean and variance of the swarm cannot be controlled simultaneously, however if the dispersion due to Brownian motion is less than the maximum controlled speed, we can adopt the hybrid, hysteresis-based controller shown in Alg.~\ref{alg:MeanVarianceControl} to regulate the mean and variance.  Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds some $\sigma_{\rm{max}}^2$. 
 Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 
 This technique satisfies control objectives that evolve at different rates as in Kloetzer's work ~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes. The process is illustrated in Fig.~\ref{fig:hysteresis}.
